### 2016-01-14 paper/master_statsoc.tex
Added keywords and changed to British spelling.

\documentclass{statsoc}
\let\proof\relax
\let\endproof\relax
\usepackage{amsmath}
\usepackage{rotating}

\newcommand{\finishcaption}{\ifx\color\undefined\else
\aftergroup\aftergroup\aftergroup
\aftergroup\aftergroup\aftergroup
\aftergroup
\aftergroup
\fi
\aftergroup\aftergroup\aftergroup
\aftergroup\aftergroup\aftergroup
\aftergroup
\leavevmode}

% Enable theorem environment
\usepackage{amsthm}
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}

\title[Asymptotics of asymmetric KDEs]{Unified treatment of the asymptotics of asymmetric kernel density estimators}

\author{Till Hoffmann}
\email{t.hoffmann13@imperial.ac.uk}
\coaddress{Department of Mathematics, Imperial College London, London SW7 2AZ, United Kingdom.}
\address{Department of Mathematics, Imperial College London, London SW7 2AZ, United Kingdom.}
\author[Hoffmann {\it et al.}]{Nick Jones}
\address{Department of Mathematics, Imperial College London, London SW7 2AZ, United Kingdom.}

\begin{document}
\input{paper}
\end{document}